### 2016-01-27 mcmc.tex
problemtweaks
two problems

% to-do
% -----
% - Deal with all mentions of ``DFM'' or ``HOGG'' and all ``CITE'' and ``...''
% - Should we cite the Brooks book or just chapters within it?  I think the latter!
% - Should we make an \acronym{} environment and use it religiously?  I think yes.
% - Do we need to explain the variance on a sampling-based estimator of an integral?  Right now this is just in an end-note (or will be when I write it).  Should it be in the main text?  Probably!
% - mention Gelman--Rubin statistic
% - mention JAGS and STAN in methods section
% - follow up the idea that burn-in doesn't exist at all!  It does for emcee.
% - when we say ``below'' or ``above'' also give a Section Number.
% - need to mention case where there is an MC integral INSIDE the likelihood function
% - need to mention testing the likelihood function
% - would love to have the Note symbols to be links to the notes and the notes to have links back to the main text...!
% - send to friendlies for comments
%   - Gelman
%   - Lang
%   - Rix
%   - Bovy
%   - Brewer
% - post on arXiv

@@

Importantly---for the algorithm given above to work correctly---the
  proposal pdf must satisfy a ``detailed balance'' condition; it
  must have the property that
\begin{eqnarray}
q(\pars'\given\pars) &=& q(\pars\given\pars')
\quad ;
\end{eqnarray}
that is, it must be just as easy to go one way in the parameter space
  as the other.
You can break this property, if you like, and then adjust the
  acceptance condition accordingly, but we \emph{do not recommend this}
  except under the supervision of a trained professional.\note{Actually,
  our own favorite method, \project{emcee}, has a proposal pdf that does
  violate detailed balance in this way and has a compensated
  acceptance ratio.  Read the paper [HOGG CITE] for more details.}
If the detailed balance condition frightens you (as it frightens us),
  then just stick with pdfs that are symmetric in $\pars'$ and $\pars$,
  like the Gaussian, or a centered uniform distribution and forget about
  it.

@@

\begin{problem}
Re-do \problemname~\ref{prob:firstMH} but with different values for
the variance of the proposal distribution $q(x'\given x)$.
What happens when you go to very extreme values (like for instance
$0.00001$ or $1000.0$)?
\end{problem}

@@

>>>>

% to-do
% -----
% - Kill all issues in this list, below:
% - Deal with all mentions of ``DFM'' or ``HOGG'' and all ``CITE'' and ``...''
% - Should we cite the Brooks book or just chapters within it?  I think the latter!
% - Should we make an \acronym{} environment and use it religiously?  I think yes.
% - Do we need to explain the variance on a sampling-based estimator of an integral?  Right now this is just in an end-note (or will be when I write it).  Should it be in the main text?  Probably!
% - mention ESJD
% - mention Gelman--Rubin statistic
% - mention JAGS and STAN in methods section
% - follow up the idea that burn-in doesn't exist at all!  It does for emcee.
% - when we say ``below'' or ``above'' also give a Section Number.
% - need to mention case where there is an MC integral INSIDE the likelihood function
% - need to mention testing the likelihood function
% - would love to have the Note symbols to be links to the notes and the notes to have links back to the main text...!
% - send to friendlies for comments
%   - Gelman
%   - Goodman
%   - Lang
%   - Rix
%   - Bovy
%   - Brewer
% - post on arXiv

@@

Importantly---for the algorithm given above to work correctly---the
  proposal pdf must satisfy a ``detailed balance'' condition; it
  must have the property that
\begin{eqnarray}
q(\pars'\given\pars) &=& q(\pars\given\pars')
\quad ;
\end{eqnarray}
that is, it must be just as easy to go one way in the parameter space
  as the other.
You can break this property, if you like, and then adjust the
  acceptance condition accordingly, but we \emph{do not recommend this}
  except under the supervision of a trained professional.\note{Actually,
  our own favorite method, \project{emcee}, has a proposal pdf that does
  violate detailed balance in this way and has a compensated
  acceptance fraction.  Read the paper [HOGG CITE] for more details.}
If the detailed balance condition frightens you (as it frightens us),
  then just stick with pdfs that are symmetric in $\pars'$ and $\pars$,
  like the Gaussian, or a centered uniform distribution and forget about
  it.

@@

\begin{problem}
Re-do \problemname~\ref{prob:firstMH} but with different values for
the variance of the proposal distribution $q(x'\given x)$.
What happens when you go to very extreme values (like for instance
$10^{-8}$ or $10^6$)?
\end{problem}

@@

\begin{problem}
Re-do \problemname~\ref{prob:firstMH} but now look at convergence:
Plot the $x$ chain as a function of timestep.  Also split the chain
into four contiguous segments (the first, second, third, and fourth
quarters of the chain).  In each of these four, compute the empirical
mean and empirical variance of $x$.  What do you conclude about
convergence from these heuristics?
\end{problem}

\begin{problem}
Write a piece of code that computes the empirical autocorrelation
function and the empirical autocorrelation time from a chain.  Run
this on the chain you obtained from \problemname~\ref{prob:firstMH}.
Plot the autocorrelation function you get and over-plot the
autocorrelation time as an indicator on the plot.  What do you
conclude?
\end{problem}
